---
title: "hw4"
author: "Ning Shen"
date: "2019/10/3"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(gapminder)
```

## Exercise 1:  Univariate Option 1

#### 1. Make a tibble with one row per year, and columns for life expectancy for two or more countries.
```{r cars}
(gapminder_wide <- gapminder %>%
   select(country, year, lifeExp) %>%
   filter(country %in% c("China", "Canada")) %>% 
   pivot_wider(id_cols = year, names_from = country, values_from = lifeExp)) %>%
   DT::datatable()
```

#### 2. Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.
```{r}
gapminder_wide %>% ggplot(aes(Canada, China)) + 
  geom_point() +
  geom_smooth(method = "lm")
```


#### 3. Re-lengthen the data.
```{r}
gapminder_wide %>%
  pivot_longer(cols = c(-year), names_to = "country", values_to = "lifeExp") %>%
  DT::datatable()
```


## Exercise 2: Multivariate Option 1

#### 1. Make a tibble with one row per year, and columns for life expectancy and GDP per capita (or two other numeric variables) for two or more countries.
```{r}
(gapminder_wide2 <- gapminder %>%
  select(year, country, lifeExp, gdpPercap) %>%
  filter(country %in% c("China", "Canada")) %>%
  pivot_wider(id_cols = year, names_from = country, names_sep = "_", values_from = c(lifeExp, gdpPercap))) %>%
  DT::datatable()
```


#### 2. Re-lengthen the data.
```{r}
gapminder_wide2 %>% 
  pivot_longer(cols = -year, names_sep = "_", names_to = c(".value", "country")) %>%
  DT::datatable()
```

## Exercise 3

```{r message=FALSE}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```


#### 3.1 For each guest in the guestlist (`guest` tibble), add a column for email address, which can be found in the `email` tibble.
```{r}
(email <- email %>% 
  rename(name = guest) %>% 
  separate_rows(name, sep = ", ")) %>%   #clean the email dataset
  DT::datatable() 

guest %>%
  left_join(email, by = "name") %>%
  DT::datatable()
```


#### 3.2 Who do we have emails for, yet are not on the guestlist?
```{r}
email %>% 
  anti_join(guest, by = "name") %>%
  DT::datatable()
```


#### 3.3 Make a guestlist that includes everyone we have emails for (in addition to those on the original guestlist).
```{r}
guest %>%
  full_join(email, by = "name") %>%
  DT::datatable()
```












